Deep Probabilistic Models for Camera Geo-Calibration

### 2018-09-05 chapters/introduction/introduction.tex
minor change
apply Brent's review opinios

The ultimate goal of image understanding is to transfer visual
signals (images/videos) into abstract symbolic descriptions of the
world which are helpful for decision making.
However, understanding images is not a trivial task for machines.
In order to handle various complicated scenarios, researchers divide
image understanding into different computer vision tasks: pedestrian
detection for autonomous driving, face recognition for security
system, and image retrieval for searching engines.

For many applications, knowing when, where, and in which direction a
picture was taken is important to understand the world. However, most
images does not carry such information. Our thesis focuses on
developing algorithms to estimating these properties. We
refer to the task as {\em geo-calibration}.

Most recent geo-calibration algorithms are deterministic systems
\todo{citations}. 
Deterministic systems have some obvious drawbacks: 1) they can not
model the inherent uncertainties from images of ambiguous scenes, and
2) without a proper modeling of the uncertainty, the subsequent
process can yield overly confident predictions. To address these
problems, we propose to build probabilistic systems for camera
geo-calibration.

We also show that learning to geo-calibrate a camera allows us to
learn about the scene of the image.
As an essential part of scene understanding, how to extract image
features is an important problem being studied for decades.  In
recent years, learning based approaches for image feature extraction
like convolutional neural networks have drawn a lot of attention due
to the fact that they do not need expert knowledge about the target
data. However, most of these feature learning methods require large
quantity of manually labeled datasets.  In this thesis, we presented
alternative ways to learn image features when the
manually labeled data is either insufficient or absent.

\section{Background}
To give a better understanding of our work, we would like to introduce
some related computer vision concepts before we discuss the main topics.
We start with the definition of geo-calibration,
along with its literature. Next, we introduce the history of deep
learning methods for geo-calibration, and a big category of approaches
that use aerial imagery. In the end, we cover some related work
of weakly supervised learning techniques.

\subsection{Geo-Calibration of Outdoor Images}
Automatic geo-calibration of outdoor images continues to grow in
importance as a direct result of the increasing amount of imagery
available via the Internet. Solving this problem is of great value
for a wide variety of fields, with potential applications ranging from
the forensic sciences~\cite{stylianou13jane} to environmental
monitoring~\cite{zhang2012mining}. Conceptually the geo-calibration
task includes identifying: the camera pose ({\em pose estimation}),
the camera geographic location ({\em geo-localization}), and the time
when the image was captured ({\em time estimation}).

@@

\item \textbf{Pose Estimation:}
Pose estimation is to estimate the camera yaw, pitch, and roll
angles, $(\theta_{yaw}, \theta_{pitch}, \theta_{roll})$ from a
image. In the context of geo-calibration, the yaw angle
$\theta_{yaw}$ is referred to as the geographic orientation angle of
the camera.
%
Li et al.~\cite{li2012worldwide} exploit geo-registered 3D points
clouds to estimate camera pose.
Vo \etal~\cite{vo2016localizing} propose a geo-localization network
that can regress the geo-orientation angle of the camera.
Agarwal \etal~\cite{agarwal2015metric} keep track of the camera pose
changes by matching SIFT~\cite{lowe1999object} feature points detected
between the input image and Google Street View panoramas.
%
Horizon line detection is an essential step for some pose
estimation algorithms, because horizon lines are closely related to
camera pitch and roll angles.
Collins and Weiss~\cite{unitsphere1990} formulate horizon line
detection as a statistical estimation problem on the Gaussian Sphere,
which is similar to the geometry we use.  More recent work has
explored the use of dual space~\cite{alignment2014,dualspace2013}
representations. Among the clustering-based approaches, Xu et
al.~\cite{kitware2013} improve this pipeline by introducing a new
point-line consistency function that models errors in the line segment
extraction step.
\newline

@@

\item \textbf{Time Estimation:}
The goal of time estimation is to estimate the time when the 
image was captured. Depending on the application, the accuracy of
predictions range from hour to year.
%
Matzen and Snavely~\cite{matzen2014scene} predict
timestamps for photos by matching against a time-varying
reconstruction of a scene.  Hill~\cite{hill1994elephant} estimate the
time of the day by measuring the light intensity profiles captured by
cameras. Methods are proposed to date the
year book photos by analyzing human
appearance~\cite{salem2016face2year,ginosar2015century}. Lee
\etal~\cite{linking2015iccp} find visual paterns in the buildings,
relate them to certain time periods.
\newline

@@

In conclusion, most of the geo-calibration algorithms we mention above
relies on finding visual cues from input images. To extract useful
visual features majority of these algorithms require strong human
expertise about the data, thus can only be applied in limited
scenarios. Therefore, we contribute to developing geo-calibration
algorithms without need of strong human expertise about the data.

\subsection{Deep Learning in Geo-Calibration}
In recent years, deep neural networks (DNNs) were proven to be
extremely successful in many computer vision areas. 
It is widely accepted that DNNs have excellent performance in
high-level visual feature learning.
This valuable ability provides researchers powerful tools to
solve the challenging geo-calibration tasks.
Weyand and Kostrikov \etal~\cite{planet} propose a convolutional
neural network (CNN) to predict the geographic location of the input
image. Walch \etal~\cite{walch2017image} aggregate learned CNN
features with LSTM to generate global image representation for camera
localization.  Studies about camera pose
estimation~\cite{zhai2016horizon, workman2016horizon,
hold2017perceptual} has been done by identifying the horizon line with
CNNs (one can derive the camera roll and pitch angles from the horizon
line position giving the camera focal length).

@@

\subsection{Ground-to-Aerial Geo-Calibration}
Data-driven scene matching approaches are commonly used in
geo-calibration.
When localizing a camera, data-driven algorithms search for k-nearest
neighbors of the query image in a database of the geo-tagged imagery.
To reduce the work of querying, the searching process is usually
carried out in feature space of applicable number of  
dimensions~\cite{im2gps, li2010location,zamir2010accurate}.
However, due to limited accuracies of GPS signals and the biased
distribution of human population, geo-tagged ground-level images
collected from smart devices are usually noisy and sparse in
geographic space. To achieve better geo-calibration results, we
need alternative image resources with more accurate geographic tags
and more complete spatial coverage.

Benefited from the fast growing monitoring satellite and drone markets,
the geo-tagged aerial imagery becomes more publicly accessible
than ever. Aerial imagery downloading services like
Microsoft Bing Map provide aerial images of various
resolutions, which cover most areas of the world with accurate
geographic registration, making aerial imagery a rich source for
reference databases. We refer to geo-calibration
methods which use the aerial imagery for geographic reference as
ground-to-aerial geo-calibration. Data used for ground-to-aerial
geo-calibration is referred to as cross-view pairs.
A cross-view pair consists of a ground image and an aerial image at
the same location, and but not necessarily aligned in orientation.

@@

The key for the success of these methods is to learn good descriptive
features for ground-to-aerial imagery matching. In the following
section, we will introduce literature about feature learning techniques.

@@

\todo{describe problems to be solve}

>>>>

The ultimate goal of image understanding is to transfer visual
signals (images/videos) into abstract symbolic descriptions of the
world which are helpful for decision making.
However, understanding images is not a trivial task for machines.
In order to handle various complicated scenarios, researchers divide
image understanding into different computer vision tasks: pedestrian
detection for autonomous driving, face recognition for security
systems, and image retrieval for search engines.

For many applications, knowing when, where, and in which direction a
picture was taken is important scene understanding. However, most
images do not carry such information. 
We refer to the task of estimating this information as {\em
geo-calibration}. Our thesis focuses on developing geo-calibration
algorithms.

Most recent geo-calibration algorithms are deterministic systems
\todo{citations}. 
Deterministic systems have some obvious drawbacks: 1) they cannot
model the inherent uncertainties from images of ambiguous scenes, and
2) without a proper modeling of the uncertainty, the subsequent
process can yield overly confident predictions. To address these
problems, we propose to build probabilistic systems for camera
geo-calibration.

We also show that learning to geo-calibrate a camera allows us to
learn about the scene of the image.
As an essential part of understanding the scene, how to extract image
features is an important problem has been studied for decades.  In
recent years, learning based approaches for image feature extraction
like convolutional neural networks have drawn a lot of attention due
to the fact that they do not need expert knowledge about the target
data. However, most of these feature learning methods require a large
quantity of manually labeled datasets. In this thesis, we propose
alternative ways to learn image features when the
manually labeled data is either insufficient or absent.

\section{Background}
To give a better understanding of our work, we would like to introduce
some related computer vision concepts before we discuss the main topics.
We start with the definition of geo-calibration,
along with its literature. Next, we introduce the history of deep
learning methods for geo-calibration, and data-driven scene matching
approaches that use aerial imagery. In the end, we cover some related
work of weakly supervised learning techniques.

\subsection{Geo-Calibration of Outdoor Images}
Automatic geo-calibration of outdoor images continues to grow in
importance as a direct result of the increasing amount of imagery
available via the Internet. Solving this problem is of great value
for a wide variety of fields, with potential applications ranging from
the forensic sciences~\cite{stylianou13jane} to environmental
monitoring~\cite{zhang2012mining}. Conceptually, the geo-calibration
task includes identifying the camera pose ({\em pose estimation}),
the camera geographic location ({\em geo-localization}), and the time
when the image was captured ({\em time estimation}).

@@

\item \textbf{Pose Estimation:}
Pose estimation is to estimate the camera yaw, pitch, and roll
angles, $(\theta_{yaw}, \theta_{pitch}, \theta_{roll})$ from an
image. In the context of geo-calibration, the yaw angle
$\theta_{yaw}$ is referred to as the geographic orientation angle of
the camera.
%
Li et al.~\cite{li2012worldwide} exploit geo-registered 3D points
clouds to estimate camera pose.
Vo \etal~\cite{vo2016localizing} propose a geo-localization network
that can regress the geo-orientation angle of the camera.
Agarwal \etal~\cite{agarwal2015metric} keep track of the camera pose
changes by matching SIFT~\cite{lowe1999object} feature points detected
between the input image and Google Street View panoramas.
%
Horizon line detection is an essential step for some pose
estimation algorithms, because horizon lines are closely related to
camera pitch and roll angles.
Collins and Weiss~\cite{unitsphere1990} formulate horizon line
detection as a statistical estimation problem on the Gaussian Sphere.
More recent work has explored the use of dual
space~\cite{alignment2014,dualspace2013} representations. Among the
clustering-based approaches, Xu et al.~\cite{kitware2013} improve this
pipeline by introducing a new point-line consistency function that
models errors in the line segment extraction step.
\newline

@@

\item \textbf{Time Estimation:}
The goal of time estimation is to estimate the time when the 
image was captured. Depending on the application, the accuracy of
predictions range from hour to year.
%
Matzen and Snavely~\cite{matzen2014scene} predict
timestamps for photos by matching against a time-varying
reconstruction of a scene.  Hill \etal~\cite{hill1994elephant} estimate the
time of the day by measuring the light intensity profiles captured by
cameras. Methods are proposed to date
yearbook photos by analyzing human
appearance~\cite{salem2016face2year,ginosar2015century}. Lee
\etal~\cite{linking2015iccp} find visual patterns in buildings,
relate them to certain time periods.
\newline

@@

In conclusion, most of the geo-calibration algorithms we mention above
rely on finding visual cues from input images. To extract useful
visual features, a majority of these algorithms require strong human
expertise about the data, thus, they can only be applied in limited
scenarios. Therefore, we contribute to developing geo-calibration
algorithms without the need of strong human expertise about the data.

\subsection{Deep Learning in Geo-Calibration}
In recent years, deep neural networks (DNNs) were proven to be
extremely successful in many computer vision areas. 
It is widely accepted that DNNs have excellent performance in
high-level visual feature learning.
This valuable ability provides researchers powerful tools to
solve the challenging geo-calibration tasks.
Weyand and Kostrikov \etal~\cite{planet} propose a convolutional
neural network (CNN) to predict the geographic location of the input
image. Walch \etal~\cite{walch2017image} aggregate learned CNN
features with LSTM to generate global image representation for camera
localization.  Studies about camera pose
estimation~\cite{zhai2016horizon, workman2016horizon,
hold2017perceptual} have been done by identifying the horizon line with
CNNs (one can derive the camera roll and pitch angles from the horizon
line position giving the camera focal length).

@@

\subsection{Ground-to-Aerial Geo-Calibration}
Data-driven scene matching approaches are commonly used in
geo-calibration.
When localizing a camera, data-driven algorithms search for k-nearest
neighbors of the query image in a database of the geo-tagged imagery.
To reduce the work of querying, the searching process is usually
carried out in a feature space of an applicable number of  
dimensions~\cite{im2gps, li2010location,zamir2010accurate}.
However, due to limited accuracies of GPS signals and the biased
distribution of human population, geo-tagged ground-level images
collected from smart devices are usually noisy and sparse in
geographic space. To achieve better geo-calibration results, we
need alternative image resources with more accurate geographic tags
and more complete spatial coverage.

Benefiting from the fast-growing monitoring satellite and drone markets,
geo-tagged aerial imagery has become more publicly accessible
than ever. Aerial imagery downloading service like
Microsoft Bing Map provides aerial images of various
resolutions, which cover most areas of the world with accurate
geographic registration, making aerial imagery a rich source for
reference databases. We refer to geo-calibration
methods which use the aerial imagery for geographic reference as
ground-to-aerial geo-calibration. Data used for ground-to-aerial
geo-calibration is referred to as cross-view pairs.
A cross-view pair consists of a ground image and an aerial image at
the same location (sometimes aligned in orientation).

@@

The key for the success of these methods is to learn descriptive
features for ground-to-aerial imagery matching. In the following
section, we will introduce literature about feature learning techniques.

@@

\todo{describe problems to solve}

### 2018-09-05 dissertation.tex
apply Brent's review opinios

\abstract
{ \SingleSpacing
The ultimate goal of image understanding is to transfer visual
images into numerical or symbolic descriptions of the scene
that are helpful for decision making.
Knowing when, where, and in which direction a picture was taken, the
task of geo-calibration, makes it possible to use imagery to understand
the world and how it changes in time. Current models for
geo-calibration are mostly deterministic, which in many cases fails to
model the inherent uncertainties when the image content is
ambiguous. Furthermore, without a proper modeling of the uncertainty, subsequent
processing can yield overly confident predictions.
To address these limitations, we propose a probabilistic model for
camera geo-calibration using deep neural networks.
While our primary contribution is geo-calibration, we also show that
learning to geo-calibrate a camera allows us to implicitly learn to
understand the content of the scene.
}
\keywords{computer vision, geo-calibration, deep neural networks}

>>>>

\abstract
{ \SingleSpacing
The ultimate goal of image understanding is to transfer visual
images into numerical or symbolic descriptions of the scene
that are helpful for decision making.
Knowing when, where, and in which direction a picture was taken, the
task of geo-calibration makes it possible to use imagery to understand
the world and how it changes in time. Current models for
geo-calibration are mostly deterministic, which in many cases fails to
model the inherent uncertainties when the image content is
ambiguous. Furthermore, without a proper modeling of the uncertainty, subsequent
processing can yield overly confident predictions.
To address these limitations, we propose a probabilistic model for
camera geo-calibration using deep neural networks.
While our primary contribution is geo-calibration, we also show that
learning to geo-calibrate a camera allows us to implicitly learn to
understand the content of the scene.
}
\keywords{computer vision, geo-calibration, deep neural networks}
